Stock enhancement undermines community stability

### 2022-02-04 rmd/figures.Rmd
edit: manuscript

**Figure 1** Theory and empirical observations agree that stock enhancement destabilizes community dynamics. In A and B, panels and colors distinguish response variables and species groups. (A) Theoretical predictions. Dots represent individual simulation replicates, and lines and shades are loess curves fitted to simulated data and 95% confidence intervals. Parameters used in this simulation are: intrinsic growth rate of the enhanced species $r_1=$ `r unique(df0$r1)`, average interspecific competition $\bar{\alpha}=$ `r unique(df0$alpha)`, carrying capacity $K=$ `r unique(df0$k)`, environmental stochasticity $\sigma_{\epsilon}=$ `r unique(df0$sd_env)`, relative fitness of captive-bred individuals $f_R=$ `r unique(df0$phi)`. (B) Empirical evidence. Dots represent geometric means of site-level observations in 31 watersheds. Lines are the predicted values of the regression models, and line types correspond to the coefficient's posterior probabilities. See Tables S8-10 for full statistics. (C) Map of sampling sites (black dots) in Hokkaido, Japan. (D) Example of a protected watershed (Okushibetsu). (E) Masu salmon *Oncorhynchus masou masou*. Photo credit: Akira Terui.

>>>>

**Figure 1** Theory and empirical observations agree that stock enhancement destabilizes community dynamics. In A and B, panels and colors distinguish response variables and species groups. (A) Theoretical predictions. Dots represent individual simulation replicates, and lines and shades are loess curves fitted to simulated data and 95% confidence intervals. Parameters used in this simulation are: intrinsic growth rate of an enhanced species $r_1=$ `r unique(df0$r1)`, average interspecific competition $\bar{\alpha}=$ `r unique(df0$alpha)`, carrying capacity $K=$ `r unique(df0$k)`, environmental stochasticity $\sigma_{\epsilon}=$ `r unique(df0$sd_env)`, relative fitness of captive-bred individuals $f_R=$ `r unique(df0$phi)`. (B) Empirical evidence. Dots represent geometric means of site-level observations in 31 watersheds. Lines are the predicted values of the regression models, and line types correspond to the coefficient's posterior probabilities. See Tables S8-10 for full statistics. (C) Map of sampling sites (black dots) in Hokkaido, Japan. (D) Example of a protected watershed (Okushibetsu). (E) Masu salmon *Oncorhynchus masou masou*. Photo credit: Akira Terui.

### 2022-02-04 rmd/manuscript.Rmd
edit: manuscript

**Affiliations:** ^1^ Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA ^2^ Salmon and Freshwater Fisheries Research Institute, Hokkaido Research Organization, Eniwa, Hokkaido 061-1433, Japan ^3^ Faculty of Environmental Earth Science, Hokkaido University, Sapporo, Hokkaido 060-0810, Japan ^4^ XXX

@@

When carrying capacity is small ($K=100$), our model predicted a destabilizing effect of stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because stocking induced the negative competitive effect exceeding the reproductive contribution of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of stock enhancement, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across most ecological contexts (robust to the changes in $r_1$, $\bar{\alpha}$, $\sigma_{\epsilon}$, and $f_R$; **Figures S1-S3**).

@@

The reduced fitness of hatchery masu salmon may also play a role in the observed response. Like other salmonids, ocean-migrating adults of hatchery masu salmon show lower return rates to the spawning river [@sahashiWildGenesBoostinpress]. In addition, hatchery salmons tend to be more susceptible to predation than wild ones [@weberInteractionsHatcheryWild2003]. The fitness disadvantage of hatchery fish may add to negative density dependence to influence community stability.

We cannot exclude the possibility that spurious correlations drove our results. However, the protected watersheds preserve nearly intact landscapes (average forest coverage: `r round(mean(df_env$frac_forest), 2) * 100`%) with strict regulations of human activities (e.g., exploitation, construction of in-stream structures); therefore, it is difficult to envision that unmeasured human influences caused the observed relationships. Further, we have statistically controlled natural variation in environmental factors (see **Materials and Methods** and **Tables S8-S10**). This unique setup may have helped uncover the close agreement between theory and empirical patterns.

There is a general agreement that the distinctive features of captive-bred individuals bring fitness disadvantages in the wild [@arakiHatcheryStockingHelp2010; @snyderLimitationsCaptiveBreeding1996]. However, despite the significant attention to the fate of captive-bred individuals, current schemes rarely consider the self-regulation process of biodiversity. Our results suggest that the ignorance of this critical process may erode the long-term persistence of the recipient community, likely impacting the provisioning of ecosystem services [@cardinaleBiodiversityLossIts2012]. While our empirical example is limited to stream fish communities, we anticipate that this phenomenon is pervasive across terrestrial and aquatic taxa because the destabilizing effect emerged across diverse simulated scenarios.

>>>>

**Affiliations:** ^1^ Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA ^2^ Salmon and Freshwater Fisheries Research Institute, Hokkaido Research Organization, Eniwa, Hokkaido, 061-1433, Japan ^3^ Faculty of Environmental Earth Science, Hokkaido University, Sapporo, Hokkaido, 060-0810, Japan ^4^ National Institute of Polar Research, Tachikawa, Tokyo, 190-8518, Japan

@@

When carrying capacity is small ($K=100$), our model predicted a destabilizing effect of stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because stocking induced the negative competitive effect exceeding the reproductive contribution of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of stock enhancement, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across most ecological contexts (robust to the changes in $r_1$, $\bar{\alpha}$, and $f_R$; **Figures S1-S3**).

@@

The reduced fitness of hatchery masu salmon may also play a role in the observed response. Like other salmonids, ocean-migrating adults of hatchery masu salmon show lower return rates to the spawning river [@sahashiWildGenesBoostinpress]. In addition, the aggressive behavior of hatchery salmons may make them vulnerable to predation [@weberInteractionsHatcheryWild2003]. The fitness disadvantage of hatchery fish may add to negative density dependence to influence community stability.

We cannot exclude the possibility that spurious correlations drove our results. However, the protected watersheds preserve nearly intact landscapes (average forest coverage: `r round(mean(df_env$frac_forest), 2) * 100`%; see also **Figure S10**) with strict regulations of human activities (e.g., exploitation, construction of in-stream structures); therefore, it is difficult to envision that unmeasured human influences caused the observed relationships. Further, we have statistically controlled natural variation in environmental factors (see **Materials and Methods** and **Tables S8-S10**). This unique setup may have helped uncover the close agreement between theory and empirical patterns.

Despite the significant attention to the fate of captive-bred individuals [@arakiHatcheryStockingHelp2010; @snyderLimitationsCaptiveBreeding1996], current schemes rarely consider the self-regulation process of biodiversity. Our results suggest that the ignorance of this critical process may erode the long-term persistence of the recipient community, likely impacting the provisioning of ecosystem services [@cardinaleBiodiversityLossIts2012]. While our empirical example is limited to stream fish communities, we anticipate that this phenomenon is pervasive across terrestrial and aquatic taxa because the destabilizing effect emerged across diverse simulated scenarios.
